---
title: "Mix9 and Conventional 16S V6 amplicons analysis"
author: "Rebecca Stevick"
date: "14 Dec 2022"
output:
  html_document:
    toc: true
    keep_md: TRUE
    theme: "cerulean"
    toc_float:
      collapsed: false
      smooth_scroll: false
---

# About the Data

## Conditions

During Trial 47. Fish collected on day 7.

-   Conventional
-   Mix9

| ID  |     SeqName     | SampleName  | Project | FishType | SampleRep |
|:---------:|:---------:|:---------:|:---------:|:---------:|:---------:|
| 71  |    71_Mix9_1    |   Mix9_1    |  Flavo  |   Mix9   |     1     |
| 72  |    72_Mix9_2    |   Mix9_2    |  Flavo  |   Mix9   |     2     |
| 73  |    73_Mix9_3    |   Mix9_3    |  Flavo  |   Mix9   |     3     |
| 74  |    74_Mix9_4    |   Mix9_4    |  Flavo  |   Mix9   |     4     |
| 75  |    75_Mix9_5    |   Mix9_5    |  Flavo  |   Mix9   |     5     |
| 76  |    76_Mix9_6    |   Mix9_6    |  Flavo  |   Mix9   |     6     |
| 77  |    77_Mix9_7    |   Mix9_7    |  Flavo  |   Mix9   |     7     |
| 78  |    78_Mix9_8    |   Mix9_8    |  Flavo  |   Mix9   |     8     |
| 79  |    79_Mix9_9    |   Mix9_9    |  Flavo  |   Mix9   |     9     |
| 80  |   80_Mix9_10    |   Mix9_10   |  Flavo  |   Mix9   |    10     |
| 86  |    86_Conv_1    |   Conv_1    |  Flavo  |   Conv   |     1     |
| 87  |    87_Conv_2    |   Conv_2    |  Flavo  |   Conv   |     2     |
| 88  |    88_Conv_3    |   Conv_3    |  Flavo  |   Conv   |     3     |
| 89  |    89_Conv_4    |   Conv_4    |  Flavo  |   Conv   |     4     |
| 90  |    90_Conv_5    |   Conv_5    |  Flavo  |   Conv   |     5     |
| 91  |    91_Conv_6    |   Conv_6    |  Flavo  |   Conv   |     6     |
| 92  |    92_Conv_7    |   Conv_7    |  Flavo  |   Conv   |     7     |
| 93  |    93_Conv_8    |   Conv_8    |  Flavo  |   Conv   |     8     |
| 94  |    94_Conv_9    |   Conv_9    |  Flavo  |   Conv   |     9     |
| 95  |   95_Conv_10    |   Conv_10   |  Flavo  |   Conv   |    10     |

## Sampling

1.  Transfer fish into a small petri dish full of sterile water (wash #1).
2.  Wash the fish again by transferring into a new petri dish full of sterile water (wash #2).
3.  Add 5 washed fish per tube into 5 x 2 mL tubes, with \~500 µL of water
4.  After all samples have been collected for all conditions, remove as much water as possible with a p200 pipette
5.  Put the samples at -80C

## Sequencing

16S V6 region

*Primers*

1064R GTCTCGTGGGCTCGGAGATGTGTATAAGAGACAGCGACRRCCATGCANCACCT\
967F-A TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGCTAACCGANGAACCTYACC\
967F-B TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGCNACGCGAAGAACCTTANC\
967F-C TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGCAACGCGMARAACCTTACC\
967F-D TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGATACGCGARGAACCTTACC

2 x 250 bp paired-end Illumina sequencing

## Analysis

Initially done with QIIME2, dada2 denoising. Eukaryotic reads removed here.

# Setup

## Load libraries and settings

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(scales)
library(ggpubr)
library(qiime2R)
library(vegan)
library(gt)
library(patchwork)
library(UpSetR)
library(gplots)
library(ComplexHeatmap)

# set global theme
theme_set(theme_minimal()+
             theme(panel.grid.major.y = element_line(color="grey80"), strip.text=element_text(size=16),
                   strip.background = element_rect(color="transparent"),
                   strip.text.y = element_text(angle=0), plot.caption = element_text(size=10),
                   panel.grid.major.x = element_blank(),legend.position="none",
                   axis.ticks = element_line(inherit.blank = FALSE),
                   panel.background = element_rect(color="grey50", size=1), 
                   legend.title = element_text(size=18),
                   plot.background = element_blank(),
                   axis.text = element_text(size=15), axis.title = element_text(size=16),
                   legend.text = element_text(size=16), plot.title = element_text(hjust=0.5)))

# set global options for code output
knitr::opts_chunk$set(echo=TRUE, warning=FALSE,message=FALSE)

```

## Import data

```{r import}

metadatafile <- "YersiniaLoperamide16Samplicon_metadata.txt"
tablefile <- "qiime2output/table.qza"
taxonomyfile <- "qiime2output/taxonomy.qza"
treefile <- "qiime2output/rooted-tree.qza"

SVs<-read_qza(tablefile)$data

metadata<-read_q2metadata(metadatafile)

taxonomy<-read_qza(taxonomyfile)$data %>% parse_taxonomy()

```

## Clean up data

```{r summarize}

# vector of eukaryotic ASVs
eukASVs <- taxonomy %>% filter(Kingdom=="d__Eukaryota" | Phylum=='Cyanobacteria') %>%
   rownames_to_column("ASVs")

datafullASVs <- t(SVs) %>% as.data.frame() %>% rownames_to_column("SampleID") %>% 
   full_join(metadata) %>% 
   pivot_longer("cca00a58137be2c40384b70e46bda58e":"213dc85b14c8cdc3e765c92bcc6fa6a2") %>% 
   # select only control Mix9/Conv data
   filter(Project=="Flavo") %>% 
   filter(LoperamidePathogen != "F. columnare") %>% 
   filter(!name %in% eukASVs$ASVs)

# ASVs <0.25% removed
datafullASVscutoff <- datafullASVs %>% group_by(SampleID) %>% 
   mutate(percent=value/sum(value)) %>% 
   filter(percent>0.001) %>% left_join(taxonomy %>% rownames_to_column("name"))

metadata <- filter(metadata, Project=="Flavo") %>% 
   filter(LoperamidePathogen != "F. columnare")
SVsControls <- select(as.data.frame(SVs), metadata$SampleID) %>%
   rownames_to_column("ASV") %>% 
   filter(ASV %in% datafullASVscutoff$name) %>% 
   column_to_rownames(var="ASV") %>% t()

```

-----

# General QC

## Reads per sample

```{r readnumber, fig.width=16, fig.height=4}

reads <-datafullASVscutoff %>% filter(Project=="Flavo") %>% 
   mutate(SampleRep=factor(SampleRep, levels=unique(SampleRep))) %>% 
   ggplot(aes(x=SampleRep, y=value, fill=Condition))+
   facet_grid(.~Condition, scales="free", space="free")+
   geom_col()+
   scale_fill_manual(values=c("red3", "gold4"))+
   scale_y_continuous(expand=c(0,0), labels=scales::label_comma())+
   labs(y="Number of reads per sample",x=NULL,fill=NULL)
reads

```


## Rarefaction curves

```{r rarefaction}

matrix <- as.data.frame(SVsControls) %>% 
   rownames_to_column("ASV") %>% 
   column_to_rownames(var="ASV")

(S <- specnumber(matrix))
(raremax <- min(rowSums(matrix)))
Srare <- vegan::rarefy(matrix, raremax)
Sslope <- vegan::rareslope(matrix, raremax)


plot(S, Srare, xlab = "Observed No. of ASVs", ylab = "Rarefied No. of ASVs")
abline(0,1)
plot(S, Sslope, xlab = "Observed No. of ASVs", ylab = "Slope at rarefied sample size")

rarecurve_data <- rarecurve(matrix, step = 100, sample = raremax)

rarecurve<-map_dfr(rarecurve_data, bind_rows) %>% 
   bind_cols(SampleID = rownames(matrix),.) %>%
   pivot_longer(-SampleID) %>%
   drop_na() %>%
   mutate(n_seqs = as.numeric(str_replace(name, "N", ""))) %>%
   select(-name) %>%
   left_join(metadata) %>% 
   ggplot(aes(x=n_seqs, y=value, group=SampleID, color=FishType, lty=FishType)) +
   geom_line(lwd=0.8) +
   scale_colour_manual(values=c("red3","gold4"))+
   scale_x_continuous(labels=label_comma())+
   theme(legend.key.width = unit(1.33,"cm"), legend.position=c(0.8,0.2))+
   labs(x = "Number of sequences", y="Number of ASVs detected", color=NULL, lty=NULL)

```

### Summary figure

```{r qcsummary, fig.width=10, fig.height=5}

( (reads|rarecurve) + plot_annotation(tag_levels = "A")) 

ggsave("figures/Mix9Conv_seqQC.png", bg="transparent", width=10, height=4)

```

----



# Taxa Bar plots


## Phylum level

```{r barsphylum, fig.width=16, fig.height=4}

palettep <- c("#D95F02", "#E6AB02","skyblue", "#1B9E77","midnightblue","grey50")

phylabars <- datafullASVscutoff %>% 
   mutate(Phylum=replace_na(Phylum, "Unknown"),
          Phylum=str_remove(Phylum, "d__Bacteria; ")) %>% 
   ggplot(aes(x=SampleName, y=value, fill=Phylum))+
   facet_grid(.~Condition, scales="free", space="free")+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.position = c(0.9, 0.6), legend.background = element_rect(fill=alpha("white", 0.8), color = "transparent"),
         axis.text.x = element_blank(), 
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"),
         axis.ticks.x = element_blank())+
   scale_fill_manual(values=palettep)+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent phyla abundance",x=NULL,fill="Phylum")
phylabars

```


## ASV level

```{r asvbarcutoff, fig.width=16, fig.height=8}

palettess<-c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99","#E31A1C", "#FDBF6F", "#FF7F00","grey40",
             "#CAB2D6", "#6A3D9A", "#FFFF99", "#B15928")

asvbars <- datafullASVscutoff %>% 
   left_join(read_qza(taxonomyfile)$data, by=c("name"="Feature.ID")) %>% 
   unite("ASVname", c("name", "Taxon"), sep=": ") %>% 
   mutate(ASVname=str_remove(ASVname, "d__Bacteria; p__"),
          TaxaOther=forcats::fct_lump_n(f=ASVname, w=value, other_level="Others", n=4),
          TaxaOther=reorder(TaxaOther, -value),
          TaxaOther=factor(TaxaOther, levels=c("4de03c7aceccf0fcb66b3589366454da: Proteobacteria; c__Gammaproteobacteria",
                              "cca00a58137be2c40384b70e46bda58e: Proteobacteria; c__Gammaproteobacteria; o__Aeromonadales; f__Aeromonadaceae; g__Aeromonas", "d3962c483c45f0765b9b7656ec7f43a0: Proteobacteria; c__Gammaproteobacteria", "d23c811ea38d6cb704bb13445f43d599: Proteobacteria; c__Gammaproteobacteria", "e1a72c614544de9ecfc0b0cd7cc24738: Bacteroidota; c__Bacteroidia; o__Flavobacteriales; f__Flavobacteriaceae; g__Flavobacterium", "33731ac0a1051571d808de3e91b11356: Proteobacteria; c__Alphaproteobacteria; o__Rhizobiales; f__Beijerinckiaceae; g__Methylobacterium-Methylorubrum; s__Methylobacterium_jeotgali", "361adfa667ee066fef8a0338a15f6305: Proteobacteria; c__Alphaproteobacteria; o__Rhizobiales; f__Rhizobiaceae; g__Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium", "b952fbdeade9c15bba8587fa6091b67f: Proteobacteria; c__Gammaproteobacteria; o__Burkholderiales; f__Comamonadaceae","Others"))) %>% 
   group_by(Condition, SampleName, TaxaOther) %>% 
   summarize(valuesum = sum(value)) %>% ungroup() %>%
   ggplot(aes(x=SampleName, y=valuesum, fill=TaxaOther))+
   facet_grid(.~Condition, scales="free", space="free")+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=12, colour="gray20"),
         legend.position = "bottom",legend.direction = "vertical", axis.text.x = element_blank(), 
         axis.ticks.x = element_blank(),
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c(palettess),
                     labels=c("4de03c7aceccf0fcb66b3589366454da: Proteobacteria; c__Gammaproteobacteria" = 
                                       "01. Proteobacteria; c__Gammaproteobacteria",
                              "cca00a58137be2c40384b70e46bda58e: Proteobacteria; c__Gammaproteobacteria; o__Aeromonadales; f__Aeromonadaceae; g__Aeromonas" = 
                                       "02. Proteobacteria; c__Gammaproteobacteria; o__Aeromonadales; f__Aeromonadaceae; g__Aeromonas",
                              "d3962c483c45f0765b9b7656ec7f43a0: Proteobacteria; c__Gammaproteobacteria" = 
                                       "03. Proteobacteria; c__Gammaproteobacteria",
                              "d23c811ea38d6cb704bb13445f43d599: Proteobacteria; c__Gammaproteobacteria" =
                                       "04. Proteobacteria; c__Gammaproteobacteria",
                              "e1a72c614544de9ecfc0b0cd7cc24738: Bacteroidota; c__Bacteroidia; o__Flavobacteriales; f__Flavobacteriaceae; g__Flavobacterium" = 
                                       "05. Bacteroidota; c__Bacteroidia; o__Flavobacteriales; f__Flavobacteriaceae; g__Flavobacterium",
                              "33731ac0a1051571d808de3e91b11356: Proteobacteria; c__Alphaproteobacteria; o__Rhizobiales; f__Beijerinckiaceae; g__Methylobacterium-Methylorubrum; s__Methylobacterium_jeotgali" = 
                                       "06. Proteobacteria; c__Alphaproteobacteria; o__Rhizobiales; f__Beijerinckiaceae; g__Methylobacterium-Methylorubrum; s__Methylobacterium_jeotgali",
                              "361adfa667ee066fef8a0338a15f6305: Proteobacteria; c__Alphaproteobacteria; o__Rhizobiales; f__Rhizobiaceae; g__Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium" = 
                                       "07. Proteobacteria; c__Alphaproteobacteria; o__Rhizobiales; f__Rhizobiaceae; g__Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium",
                              "b952fbdeade9c15bba8587fa6091b67f: Proteobacteria; c__Gammaproteobacteria; o__Burkholderiales; f__Comamonadaceae" = 
                                       "08. Proteobacteria; c__Gammaproteobacteria; o__Burkholderiales; f__Comamonadaceae",
                              "Others" = "Others"))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent ASV abundance", x=NULL, fill="ASV")
asvbars

```

### Summary figure

```{r combinebars, fig.width=13, fig.height=12}

phylabars/asvbars+plot_annotation(tag_levels = "A")+
  theme(plot.tag = element_text(size=20))
ggsave("figures/Mix9ConvBars.png", bg="transparent", width=13, height=11)

```


-----

# Venn Diagrams

At the ASV level

```{r venn, fig.width=10}

convset <- datafullASVscutoff %>% ungroup() %>% filter(Condition=="Conv") %>% distinct(name) %>% .$name
mix9set <- datafullASVscutoff %>% ungroup() %>% filter(Condition=="Mix9") %>% distinct(name) %>% .$name

venn(list("Conv" = convset, "Mix9" = mix9set))

vennd <- ggVennDiagram::ggVennDiagram(list("Conv" = convset, "Mix9" = mix9set))+
   scale_color_manual(values=c('red3','gold4'))+
   scale_fill_gradient(low="white",high="grey50")+
   coord_sf(clip = 'off')+theme(legend.position = "none")
vennd

```



-----

# Alpha Diversity

## Richness chao

```{r alphachao}

chao <- estimateR(SVsControls) %>% t() %>% as.data.frame()
metadata$Chao <- chao$S.chao1

# calculate diversity
metadata %>% 
   ggplot(aes(x=Condition,y=Chao, fill=Condition, shape=Condition))+
   geom_jitter(width=0.15, size=3, alpha=0.8)+
   geom_boxplot(alpha=0.8)+
   stat_compare_means(ref.group="Conv", label = "p.format", vjust=1)+
   labs(title="Chao Index", x=NULL,y="Chao Index",fill="Condition")+
   scale_fill_manual(values=c("red3","gold4")) +
   scale_shape_manual(values=c(22,23))
 
```

## Simpsons and Shannon

```{r alphasimpshan, fig.width=8, fig.height=5}

# calculate diversity
diversityshannon<-diversity(SVsControls, index="shannon")
diversitysimpsons<-diversity(SVsControls, index="simpson")
# add into metadata variable
metadata$Simpsons<-diversitysimpsons
metadata$Shannon<-diversityshannon

divplot <- metadata %>% 
   pivot_longer(Chao:Shannon, names_to="DiversityIndex", values_to="Value") %>% 
   ggplot(aes(x=Condition,y=Value, fill=Condition, shape=Condition))+
   geom_jitter(width=0.15, size=3, alpha=0.8)+
   geom_boxplot(alpha=0.8)+
   facet_wrap(~DiversityIndex, scales="free")+
   stat_compare_means(ref.group="Conv", label = "p.format")+
   labs(x=NULL, y="Index of Diversity",fill="LoperamideTimepoint")+
   scale_fill_manual(values=c("red3","gold4")) +
   scale_shape_manual(values=c(22,23))+
   scale_y_continuous(expand = expansion(mult = c(0.1, .1)), limits=c(0,NA))
divplot 

```


-----


# Beta Diversity

```{r betadivsetup}

# ellipse function
veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100)
{theta <- (0:npoints) * 2 * pi/npoints
   Circle <- cbind(cos(theta), sin(theta))
   t(center + scale * t(Circle %*% chol(cov)))}

```

```{r convmix9alltogether,  fig.show='hide', fig.width=4}

sol<-metaMDS(SVsControls,distance = "bray", k = 2, trymax = 200)
sol$stress; stressplot(sol)

NMDS=data.frame(x=sol$point[,1],y=sol$point[,2],
                FishType=as.factor(metadata$FishType),
                Rep=as.factor(metadata$SampleRep))
plot.new()
ord<-ordiellipse(sol, NMDS$FishType, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$FishType)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$FishType==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),FishType=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$FishType),mean)
```

```{r plotmix9convbeta}
betadiv<-ggplot(data=NMDS,aes(x,y,colour=FishType,fill=FishType))+
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=FishType), size=1) +
  geom_point(size=4, alpha=0.8, aes(shape=FishType), color="black")+
   scale_shape_manual(values=c(22,23))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") +
  scale_fill_manual(values=c("red3","gold4"))+scale_colour_manual(values=c("red3","gold4"))+
   ggtitle("Bray-Curtis beta-diversity")
betadiv
```


## Within group beta-diversity
```{r withinbeta}

braycurtisdistances<-
  # calculate the dissimilarity matrix between each sample
  vegdist(SVsControls, method="bray", k=2) %>% as.matrix() %>% as.data.frame() %>% 
  # Add in the sites based on the rows (paired sample)
  rownames_to_column(var="PairedID")  %>% mutate(PairedCondition=metadata$Condition)%>%
  # Make into longform based on the columns (OG SampleID)
  pivot_longer(cols="71-Mix9-1":"95-Conv-10", names_to="SampleID") %>% 
  # Add in sites based on the columns (OG SampleID)
  left_join(metadata) %>% 
  # Remove rows where a sample is paired with a sample from another condition
  filter(PairedCondition==Condition) %>% 
  # Remove the diagonal rows, where each sample was compared to itself
  filter(value!=0)

# stats
compare_means(data=braycurtisdistances, value~Condition, method="wilcox", p.adjust.method = "BH")
compare_means(data=braycurtisdistances, value~Condition, method="kruskal")

# plot within condition variation
withinbeta<-ggplot(braycurtisdistances, aes(x=Condition, y=value, fill=Condition, shape=Condition))+
  geom_jitter(width=0.15, size=2,  alpha=0.7)+geom_boxplot(alpha=0.8)+
  labs(x=NULL,y="within condition \ndissimilarity index",fill="Condition")+
   stat_compare_means(ref.group="Conv", label = "p.format", vjust=1)+
   scale_fill_manual(values=c("red3","gold4")) +
   scale_shape_manual(values=c(22,23))+
  scale_y_continuous(limits=c(0,1), labels=c("0.00","0.25","0.50","0.75","1.00"))
withinbeta

```

## Summary figure

```{r plotalldiv, fig.width=10, fig.height=10}

((vennd+plot_spacer())) / divplot / (betadiv+withinbeta+plot_layout(widths=c(1.8,1))) +
   plot_annotation(tag_levels = "A")+plot_layout(heights=c(4,4,4))

ggsave("figures/DiversitySummary.png", bg="transparent", width=7, height=8)

```


```{r sessioninfo}
sessionInfo()
```

